---
title: "HW5"
author: "Vincent Roberson"
date: today
format: pdf
execute:
  echo: true
---

# HW 05- Introduction (100 pts).
## Assignment Overview
The purpose of this assignment is to review maximum likelihood estimators (MLE) and practice solving MLE analytically and computationally.  

## Instructions
- You will answer the questions directly in the Rmarkdown template file provided. 
- Please change your name in the template file.
- How to insert R code chunks: You can quickly insert chunks into your R Markdown file with the keyboard shortcut Cmd + Option + I (Windows Ctrl + Alt + I).
- You can knit a PDF and upload the final pdf file to Canvas.


## Part I Derive the MLE for the parameters in the following distributions (Total 50 pts, 25 pts each)
(note: You may take a picture of what you have done on a paper and insert your pictures into Rmarkdown or type the solution in LaTeX format under each question. )

A. The Poisson distribution is often used to describe how many times an event takes place within a certain time interval, such as number of daily customers at a store. Assume $x_1, \ldots, x_n$ independently drawn from Poisson distribution ($\lambda)$, where $\lambda$ is the rate of occurrences. Derive the maximum likelihood estimator of $\lambda$ . The
  probability mass function (PMF) for the Poisson distribution is:

 $f(x | \lambda) = \frac{\lambda^x e^{-\lambda}}{x!}$




B. Bernoulli distribution is a special binomial distribution when the parameter n = 1. The PDF for Bernoulli distribution is $f(x | p) = p^x(1-p)^{1-x}$. Please derive the MLE for $p$ given n observations $x_1, \ldots, x_n$ independently and identically drawn from the Bernoulli distribution with parameter $p$.



## Part II: Computational optimization for getting MLE using `optim` function in R. (50 pts)

A. Assume $x_1, \ldots, x_N$ are independently drawn from Poisson($\lambda$).  Using the simplified log-likelihood for the Poisson distribution from Part I A.

a) Generate 1000 random numbers from Poisson distribution with parameter $\lambda = 10$. 

``` {r}
rpois(n = 1000, lambda = 10) -> random_numbers
```

b) Write the simplified log-likelihood in R code (Please show your code).

``` {r}
poisson_log_likelihood <- function(lambda, x) {
  sum(x * log(lambda) - lambda - lfactorial(x))
}
```
	
c) Write `R` code using the function `optim` to find the maximum likelihood estimator for $\lambda$. 

``` {r}
poisson_max_likelihood_estimator <- optim(
  par = 1,
  fn = poisson_log_likelihood,
  method = "L-BFGS-B", lower = 0.01,
  control = list(fnscale = -1),
  x = random_numbers
)$par
poisson_max_likelihood_estimator
```


B. Assume $x_1, \ldots, x_N$ are independently drawn from Bernoulli($p$).  Using the simplified log-likelihood for the Bernoulli distribution from Part I B.

a) Generate 1000 random numbers from Bernoulli distribution with parameter $p = 0.15$. 

``` {r}
bernouli_random_numbers <- rbinom(1000, size = 1, prob = 0.15)
```

b) Write the simplified log-likelihood in R code (Please show your code).

``` {r}
bernoulli_log_likelihood <- function(p, x) {
  sum(x*log(p) + (1 - x) * log(1 - p))
}
```
	
c) Write `R` code using the function `optim` to find the maximum likelihood estimator for $p$. 

``` {r}
bernoulli_log_likelihood_estimator <- optim(
  par = 0.5,
  fn = bernoulli_log_likelihood,
  method = "L-BFGS-B", lower = 0.01, upper = .99,
  control = list(fnscale = -1),
  x = bernouli_random_numbers
)$par
bernoulli_log_likelihood_estimator
```


## Part III Bonus (Optional, the purpose is to encourage your efforts in learning and character growth. Bonus points could be given but not limited to the following efforts)
a. Attend community coding or office hours: please list the date, time and the faculty/student assistant you met.

b. Write a statement of how you overcome the difficulties in life in order to learn.

In order to overcome the difficulties in life in order to learn, I will make sure that I complete all my work for the day and portion of the work due farther along the long at a consistent rate in order to avoid binge sessions caused by external life commitments where learning ability is lessoned.

c. Write your questions regarding lectures. 






